% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{mathtools}
\usepackage{braket}
\usepackage{csquotes}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={The Meßthaler-Wulff Project},
  pdfauthor={Julia Meßthaler},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{The Meßthaler-Wulff Project}
\author{Julia Meßthaler}
\date{}

\begin{document}
\maketitle

Blazingly fast code for finding all crystals (subsets of a graph) that
can be constructed using only transformations that locally minimize
surface energy.

\hypertarget{the-problem}{%
\subsection{The Problem}\label{the-problem}}

Let \((G, E)\) be some graph and \(\mathcal{N}: G \rightarrow \wp(G)\)
denote the neighbors of a given node, defined as \[
  \mathcal{N}(n) \coloneq \{ n_0 \in G \mid \{n_0, n\} \in E \}
\]

Now we can define a crystal as \(c \subset G\) with \(c\) finite. This
allows us to define the set of all crystals
\(C \coloneq \{ c \subset G \mid c \text{ finite} \}\).

Now we can define the surface energy of the crystal \(c\) (or arbitrary
subsets of \(G\)) as \[
  E_c \coloneq \sum_{n \in c} f_{G \setminus c}(n)
\] where \(f_M(n)\) denotes the \textquote{friendliness} of the node
\(n\) or how many friends it has defined as \[
  f_M(n) \coloneq \# \{ n_0 \in \mathcal{N}(n) \mid n_0 \in M \}
\]

The idea now is to find crystals \(c\) such that \(\frac{E_{c}}{\#c}\)
is optimal.

\hypertarget{the-crystal-graph}{%
\subsection{The Crystal Graph}\label{the-crystal-graph}}

We can impose a graph structure on \(C\) with the edges \[
  T \coloneq \{ \{ c, c \setminus \{n\} \} \mid c \in C \text{ and } n \in c \}
\] we call these the transformations and \((C,T)\) the transformation
graph.

If we want to discover optimal crystals, then we must efficiently walk
this graph. Since this graph is very dense and very large, we must first
discuss some optimizations:

\begin{itemize}
\tightlist
\item
  \protect\hyperlink{efficient-transformations}{\(O(1)\)
  transformations}
\item
  \protect\hyperlink{pruning-bad-transformations}{Pruning bad
  transformations}
\item
  \protect\hyperlink{exploiting-symmetries}{Exploiting Symmetries}
\item
  \protect\hyperlink{pruning-bad-crystals}{Pruning bad crystals}
\end{itemize}

\hypertarget{efficient-transformations}{%
\subsection{Efficient Transformations}\label{efficient-transformations}}

In the code this is achieved using the stateful class
\texttt{AdditiveSimulation} that walks along the transformation graph.
This class keeps track of two important properties, namely \(f_c(n)\)
and \(\chi_c(n)\) for (almost) all \(n \in G\), we do not need to store
the (possibly) infinite number of values, as we can let \(f_c(n)\)
default to \(0\) and \(\chi_c(n)\) to \(0\). \(\chi_c(n) \in \{0,1\}\)
is the characteristic function of the set \(c\) and is equal to \(1\)
exactly iff the node is in \(c\).

To the characteristic function we can associate a sign using
\(s: \{0,1\} \rightarrow \{-1, 1\}\), defined as \(s(x) = 2x - 1\).

Now let us take a look at what happens, if we walk along an edge
\(t \in T\). Let \(n\) be the single node that is affected by this
transformation.

Let \(\varphi\) be any specific integer quantity we are tracking, then
let \(\Delta\varphi \coloneq \varphi' - \varphi\) be the difference
between the old and the new value of the quantity.

\(\chi_c\) is updated using \(\Delta\chi_c(n) = -s\circ\chi_c(n)\).

The friendliness \(f_c(n)\) does not change. Instead, the friendliness
of each neighbor must be updated. So for each
\(n_0 \in \mathcal{N}(n)\): \(\Delta f_c(n_0) = -s\circ\chi_c(n)\).

For updating the Energy \(E_c\) will first require some work. We already
know the update rule for neighbors \(n_0\) of \(n\). For nodes \(n_0\)
that are not \(n\) or neighbors of \(n\) the differential is
\(\Delta f_c(n_0) = 0\), in essence this is also not problematic.

However for \(n\) the differential is more involved, as the transition
makes the node become part of or be excluded from the sum that defines
\(E_c\). So its contribution is \[
    -s\circ\chi_c(n) \cdot f_{G \setminus c}(n) = -s\circ\chi_c(n) \cdot [\#\mathcal{N}(n) - f_c(n)]
\]

Now for the calculation: \[
    \Delta E_c   = \sum_{n_0 \in c} f_{G \setminus c}(n_0)
\]

\[
            = \left[ \sum_{n_0 \in \mathcal{N}(n) \cap c} \Delta f_{G \setminus c}(n_0)\right] -s\circ\chi_c(n) \cdot [\#\mathcal{N}(n) - f_c(n)]
\]

\[
            = \left[ \sum_{n_0 \in \mathcal{N}(n) \cap c} s\circ\chi_c(n) \right] -s\circ\chi_c(n) \cdot [\#\mathcal{N}(n) - f_c(n)]
\]

\[
            = s\circ\chi_c(n) \cdot f_c(n) -s\circ\chi_c(n) \cdot [\#\mathcal{N}(n) - f_c(n)]
\]

\[
            = s\circ\chi_c(n) \cdot (2f_c(n) - \#\mathcal{N}(n))
\]

\hypertarget{pruning-bad-transformations}{%
\subsection{Pruning Bad
Transformations}\label{pruning-bad-transformations}}

Most transformations are pretty bad, for example in a solid crystal,
removing a node in the center will only increase the energy, to combat
this we will only consider locally optimal transformations.

This means that we only take transformations \(t\) for which
\(\Delta E_c\) is at most as large as the \(\Delta E_c\) for all other
transformations from this node \(c\).

We can do this by directly storing \(\Delta E_c\) for every node or a
similar proxy. Luckily we are already storing \(f_c\) and we can simply
derive \(\Delta E_c\).

Now, we do not just need to know the energy differential, but we also
need to be able to query all nodes that have this minimal differential.

This is achieved in \(O(1)\) using the \texttt{PriorityStack} class, a
priority queue optimized for this specific use case. We have two
instances for our graph walking \texttt{AdditiveSimulation}, one for
nodes that are not in the crystal and one for the ones that are.
Depending on which direction is then queried (addition or removal of
nodes), we query the appropriate instance for all nodes that have
minimal/maximal \(f_c(n)\).

\hypertarget{exploiting-symmetries}{%
\subsection{Exploiting Symmetries}\label{exploiting-symmetries}}

Let \(H\) be a group action on \(G\). This is only useful to us if the
action commutes with the neighborhood function as in
\(\eta(h(n)) = h(\eta(n))\) for all \(h \in H\). This not only means
that the energy is also invariant under \(H\), but also the possible
next locally optimal transformations, meaning we can run our simulation
only on canonical representatives of each equivalence class of \(G/H\).

We call a function \(\chi_H: G \to H\) a characteristic for the group
action \(H\) if for all \(h \in H\) and \(g \in G\)
\[\chi_H(h(g)) = h \circ \chi_H(g)\]

\hypertarget{pruning-bad-crystals}{%
\subsection{Pruning Bad Crystals}\label{pruning-bad-crystals}}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\end{document}
